---
title: "Lab 04 - La Quinta is Spanish for next to Denny's, Pt. 2"
author: "Ben Hardin"
date: "2/7/2023"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
dennys <- dennys
laquinta <- laquinta
```

### Exercise 1

First, let's make a dataframe that only has Denny's locations in Alaska. Then, we can get the number of rows, which shows that there are 3 Denny's in Alaska.

```{r dennys-alaska}
dennys_ak <- dennys %>%
  filter(state == "AK")
nrow(dennys_ak)
```

### Exercise 2

Now, let's do the same for LaQuinta locations. It seems there are 2 Laquintas in Alaska.

```{r laquinta-alaska}
laquinta_ak <- laquinta %>%
  filter(state == "AK")
nrow(laquinta_ak)
```

### Exercise 3

There are 3 Denny's and 2 LaQuintas, so there are 3*2 = 6 distances between Denny's and LaQuinta locations in Alaska. Let's join the two datasets we created, so we can calculate these distances.

```{r joining-forces}
dn_lq_ak <- full_join(dennys_ak, laquinta_ak, by = "state")
dn_lq_ak
```

### Exercise 4

There are 6 rows and 11 columns in the new dataset. The rows represent the pairings of Denny's and LaQuinta locations in Alaska. The columns represent the variables. There is only one variable for state, because that is the variable we joined the Denny's and LaQuinta datasets on. Then, we have two copies of every other location variable, with an added letter (either x or y), which tells us whether the observation came from the Denny's dataset (x) or Laquinta dataset (y)

### Exercise 5

The mutate function in tidyverse lets us add a new variable to our dataset.

```{r inventing-haversine-function}
haversine <- function(long1, lat1, long2, lat2, round = 3) {
  # convert to radians
  long1 = long1 * pi / 180
  lat1  = lat1  * pi / 180
  long2 = long2 * pi / 180
  lat2  = lat2  * pi / 180
  
  R = 6371 # Earth mean radius in km
  
  a = sin((lat2 - lat1)/2)^2 + cos(lat1) * cos(lat2) * sin((long2 - long1)/2)^2
  d = R * 2 * asin(sqrt(a))
  
  return( round(d,round) ) # distance in km
}
```

### Exercise 6

Alrighty, now that we have the two functions we need to do so, let's calculate the distances between each Denny's and LaQuinta.

```{r distance-variable}
dn_lq_ak <- dn_lq_ak %>%
  mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y))
```

### Exercise 7

Now we can calculate the minimum distance of a LaQuinta from each Denny's. 

```{r min-distance}
ak_mindist <- dn_lq_ak %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))
```

### Exercise 8

Let's summarize the distance between each Denny's in Alaska and each one's nearest LaQuinta Inn. The average distance from a given Denny's to a LaQuinta Inn is 4.4. For any given Denny's in Alaska, you are at most about 6 km away from LaQuinta Inn; and if you're at the luckiest Denny's of all, you are only 2 km away from a LaQuinta. 

I made a histogram to visualize the distribution of distances from each Denny's and its nearest LaQuinta, rendered in beautiful Denny's gold! It's not too exciting, because there are only 3 locations...

```{r summarize-distances}
ak_mindist %>%
  summary(closest)

#getting factors ordered by closeness to LaQuinta
ak_mindist <- ak_mindist %>%
  arrange(closest) %>%
  mutate(address.x = factor(address.x, levels = address.x))
  
#plotting
ggplot(ak_mindist, aes(x = closest))+
  geom_histogram(color = "brown", fill = "gold")+
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 2))+
  theme_bw()+
  labs(title = "Distribution of Denny's to LaQuinta distances in Alaska", y = "Count", x = "Distance between each Denny's and it's the nearest LaQuinta (km)")
```

### Exercise 9

Now, we repeat the whole process for North Carolina. There's a nicer selection of locations now, and we can see that there's quite a spread, where some Denny's have a LaQuinta very nearby, but a substantial number of Denny's in NC are more than 100 km away from a LaQuinta.

```{r dennys-nc}
dennys_nc <- dennys %>%
  filter(state == "NC")
nrow(dennys_nc)
```

```{r laquinta-nc}
laquinta_nc <- laquinta %>%
  filter(state == "NC")
nrow(laquinta_nc)
```

```{r joining-nc}
dn_lq_nc <- full_join(dennys_nc, laquinta_nc, by = "state")
dn_lq_nc
```


```{r distance-nc}
dn_lq_nc <- dn_lq_nc %>%
  mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y))
```


```{r min-nc}
nc_mindist <- dn_lq_nc %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))
```


```{r plotting-nc}
nc_mindist <- nc_mindist %>%
  arrange(closest) %>%
  mutate(address.x = factor(address.x, levels = address.x))

ggplot(nc_mindist, aes(x = closest))+
  geom_histogram(color = "brown", fill = "gold")+
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 3.5))+
  theme_bw()+
  labs(title = "Distribution of Denny's to LaQuinta distances in North Carolina", y = "Count", x = "Distance between each Denny's and it's the nearest LaQuinta (km)")
```

### Exercise 10

Now, we repeat the whole process for Texas. A couple of things are striking. First, there are a ton of Denny's in Texas! Second, a huge chunk of the Denny's in Texas are only a few kilometers away from a LaQuinta, and only a tiny few Denny's are more than 20 km from a LaQuinta.

```{r dennys-tx}
dennys_tx <- dennys %>%
  filter(state == "TX")
nrow(dennys_tx)
```

```{r laquinta-tx}
laquinta_tx <- laquinta %>%
  filter(state == "TX")
nrow(laquinta_tx)
```

```{r joining-tx}
dn_lq_tx <- full_join(dennys_tx, laquinta_tx, by = "state")
dn_lq_tx
```


```{r distance-tx}
dn_lq_tx <- dn_lq_tx %>%
  mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y))
```


```{r min-tx}
tx_mindist <- dn_lq_tx %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))
```


```{r plotting-tx}
tx_mindist <- tx_mindist %>%
  arrange(closest) %>%
  mutate(address.x = factor(address.x, levels = address.x))

ggplot(tx_mindist, aes(x = closest))+
  geom_histogram(binwidth = 1, color = "brown", fill = "gold")+
  scale_y_continuous(
    expand = c(0, 0),
    limits = c(0, 40))+
  theme_bw()+
  labs(title = "Distribution of Denny's to LaQuinta distances in Texas", y = "Count", x = "Distance between each Denny's and it's the nearest LaQuinta (km)")
```

### Exercise 11

Okay, now let's try it one more time with my home state, Mississippi. This one is a little bit like Alaska, because there aren't very many Denny's in Mississippi. Of those, it looks like a couple of them are quite close to a LaQuinta, one of them is about 20 km away, and the last one is more than 100 km away.

```{r dennys-ms}
dennys_ms <- dennys %>%
  filter(state == "MS")
nrow(dennys_ms)
```

```{r laquinta-ms}
laquinta_ms <- laquinta %>%
  filter(state == "MS")
nrow(laquinta_ms)
```

```{r joining-ms}
dn_lq_ms <- full_join(dennys_ms, laquinta_ms, by = "state")
dn_lq_ms
```


```{r distance-ms}
dn_lq_ms <- dn_lq_ms %>%
  mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y))
```


```{r min-ms}
ms_mindist <- dn_lq_ms %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))
```


```{r plotting-ms}
ms_mindist <- ms_mindist %>%
  arrange(closest) %>%
  mutate(address.x = factor(address.x, levels = address.x))

ggplot(ms_mindist, aes(x = closest))+
  geom_histogram(color = "brown", fill = "gold")+
  scale_y_continuous(
    expand = c(0, 0),
    limit = c(0, 2.1))+
  theme_bw()+
  labs(title = "Distribution of Denny's to LaQuinta distances in Mississippi", y = "Count", x = "Distance between each Denny's and it's the nearest LaQuinta")
```

### Exercise 12

Mitch Hedberg's joke seems to hold the most true for Texas, because so many Denny's are within 5 km of the nearest LaQuinta! Alaska is a close second, except there are so few Denny's there that it seems kind of irrelevant.